DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-08-02 2015CompleteDerivation/completeDerivation.tex
Reviewed language for classical assumption

	The second, more conceptual, problem is that the idealizations that allow us to hold Assumption \ref{detrevass} break down as the parts get smaller. One such idealization is when the system under study is "big enough". Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface. Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and therefore not sufficiently isolated. The other idealization is when the interaction between system and environment is not symmetrical. For example, suppose the electric charge of the system under study is much larger than the charges in the environment. The electromagnetic field generated can be considered determined only by our system with good approximation. But this also breaks down: if each part of the environment is made of the same homogeneous material the charge density will be the same and the interaction between particles will be symmetrical.

	In other words: the definition of a state requires some kind of asymmetry between system under study and environment. The environment should not affect the system, so that we can define an independent state, yet the system must affect the environment, so that we have external process that allows to define physical distinguishability. If we try to assume that both the system and environment are really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost. Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment). For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average motion is sufficiently stable.

	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we want the environment to not be isolated so that we can define physical processes that interact with the system. The problem is that the same physical object could be either under study or part of the environment depending on the situation. In other words: we want systems to be interacting so that they can be physically aware of each other, yet we want them not interacting so that we can properly define states on them. We clearly can't have it both ways. Under our classical assumption, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts. Each particle would essentially reside in its own separate physical universe, as it cannot physically distinguish anything else. This cannot work.

	If we assumed the system is deterministic and reversible only as a whole, then each part could evolve depending on the states of the other parts. This would seem much better, as we have pieces with no unstated part that are physically aware of each others. But this does not actually solve the problem of independent oblivious components. Let's ignore the question of how we were able to define the states of the parts. Let's also ignore that the type of forces one would typically use (i.e. ones that conserve energy or ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other. The actual problem is that we can always locally separate the evolution into independent degrees of freedom. For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently. And since the pieces are infinitesimal, a local decomposition is all that is needed. While such description may be cumbersome to achieve in practice, conceptually it is still possible. So, even if the particles are in principle interacting, the system is described by degrees of freedom that evolve independently, oblivious to each other.

	The same type of problem arises on the whole system when we consider its interaction with our probe or other parts of the environment. Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system. This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map. In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems if the parts interact in any physically meaningful way.

	The moral is that we should \emph{always} beware of thinking that the state of a system is its full description: there is always some unstated part that allows continuous interaction between the system and the environment. That is the part that cannot be isolated (i.e. we can't take the system out of the physical universe) and therefore can't be assigned a state.

	One final conceptual problem is with time itself: a deterministic and reversible system cannot tell time. If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes. The problem is that deterministic and reversible time evolution is equivalent to a state variable change. In fact, at a fixed time, we can choose to describe the system with the transported state variables of any past or future times. So, all quantities that are invariant under state variable changes are also invariant under time evolution.

	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to strictly apply to all of universe. While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts. Understanding its shortcomings will help us later to see how the quantum assumption solves, at least partially, some of these issues.

>>>>

	The second, more conceptual, problem is that the idealizations that allowed us to define states in Assumption \ref{detrevass} may not hold as the parts get smaller. Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface. As we consider smaller and smaller parts, at some point we will find an amount of material that is affected by the interaction with the air or photons scattering off of it. At that point the parts are no longer sufficiently isolated to define an independent state, their evolution depends on the environment. Therefore the deterministic evolution of the whole cannot be reduced to the deterministic evolution of its parts.

	%The other idealization is when the interaction between system and environment is not symmetrical. For example, suppose the electric charge of the system under study is much larger than the charges in the environment. The electromagnetic field generated can be considered determined only by our system with good approximation. But this also breaks down: if each part of the environment is made of the same homogeneous material the charge density will be the same and the interaction between particles will be symmetrical.

	Another issues is that defining a state requires some kind of asymmetry between system under study and environment. The system evolution is not affected by the state of the environment, so that we can define an independent state, yet the environment evolution is affected by the state of the system, so that we have external processes that allow us to define physical distinguishability. If we try to assume that both the system and environment are really made of the same classical material, then the claim to that asymmetry is lost: there is no difference in size, everything is made of infinitesimal pieces, and there is no difference in the laws of evolution, as the pieces are homogeneous.

	Another issue arises if we include the probe in our description. Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system. This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map. If we want determinisms and reversibility for the combined system, we either have two isolated systems $\hat{a}=f(a)$ and $\hat{b}=g(b)$, which means the probe can't learn anything about the system, or two interacting systems $\hat{a}=f(a, b)$ and $\hat{b}=g(a, b)$, which means our system is not deterministic as its future state is determined by the state of the probe. In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregations and have the parts interact in any physically meaningful way.

	%Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment). For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average motion is sufficiently stable.

	%Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we want the environment to not be isolated so that we can define physical processes that interact with the system. The problem is that the same physical object could be either under study or part of the environment depending on the situation. In other words: we want systems to be interacting so that they can be physically aware of each other, yet we want them not interacting so that we can properly define states on them. We clearly can't have it both ways.

	The same problem of not meaningful interaction surfaces when we consider the states of infinitesimal parts. Under our classical assumption, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts. Each particle would essentially reside in its own separate physical universe, as it cannot physically distinguish anything else. This is not physically meaningful.

	If we assumed the system is deterministic and reversible only as a whole, then each part could evolve depending on the states of the other parts. This would seem much better, as the state of the pieces is still exhaustive yet they are physically connected to each others. But this does not actually solve the problem of independent oblivious components. First, for this to work, one would have to specify how the states of the parts were defined since their evolution is no longer deterministic (the future of each part depends on the state of other parts). One would also need to explain what type of interaction exists between the parts: simply using conservative forces, for example, would smell of a circular argument since those are precisely the force that we derive in the case where the pieces do evolve independently. But ignoring these two issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom. For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently. And since the pieces are infinitesimal, a local decomposition is all that is needed. While such description may be cumbersome to achieve in practice, conceptually it is still possible. So, even if each particle evolution depends on the state of the others, the system is described by degrees of freedom that evolve independently, oblivious to each other.

	The moral is that the classical idea of being able to assign to all systems a state which represents its full description does not work. The state is only what we can describe and it can't be the full description. As the division between system and environment is subjective, each system must be able to function as both. Therefore it will have a part whose evolution depends only on the system itself, the state, and a part whole evolution depends on other systems, the unstated part. While they may not be the same in all circumstances, they must exist. The state is what gives the identity to the system, its the subject of our discussion. The unstated part is what allows continuous interaction between the system and the environment, its what allows us to have the discussion. As we'll see later, it is precisely this problem that quantum system addresses better conceptually than classical mechanics.

	One final problem is with time itself: an infinitesimally reducible deterministic and reversible system cannot tell time. If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes. The problem is that deterministic and reversible time evolution is equivalent to a state variable change. In fact, at a fixed time, we can choose to describe the system with the transported state variables of any past or future times. So, all quantities that are invariant under state variable changes are also invariant under time evolution.

	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to strictly apply to all of the universe. While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts. Understanding its shortcomings will help us later to see how the quantum assumption solves, at least partially, some of these issues.
